### 2022-03-09 paper/output/document/thesis_0_abstract.tex
making my way dowtown, typing fast, appendices pass

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}abstract}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
AI agents are beginning to make decisions without human supervision in increasingly consequential 
contexts like healthcare, policing, and driving. These decisions are inevitably ethically tinged, 
but most AI agents navigating the world today are not explicitly guided by ethics.
Regulators, philosophers, and computer scientists are raising the alarm about the 
dangers of unethical artificial intelligence, from science-fiction killer robots to criminal
sentencing algorithms prejudiced against people of color. This is spurring interest in automated ethics, or the development 
of machines that can perform ethical reasoning. Much prior work in automated ethics approaches the 
problem from a computational perspective and rarely engages with philosophical literature on ethics, despite
its relevance to the development of AI agents that can
responsibly navigate the world. If automated ethics draws on sophisticated philosophical literature, the ethical reasoning
underlying such decisions will be more nuanced, precise, consistent, and trustworthy. However, faithfully translating complex ethical theories
from natural language to the rigid syntax of a computer program poses technical and philosophical 
challenges.

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}abstract}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
AI agents are beginning to make decisions without human supervision in increasingly consequential 
contexts like healthcare, policing, and driving. These decisions are inevitably ethically tinged, 
but most AI agents navigating the world today are not explicitly guided by ethics.
Regulators, philosophers, and computer scientists are raising the alarm about the 
dangers of unethical artificial intelligence, from lethal autonomous weapons to criminal
sentencing algorithms prejudiced against people of color. This is spurring interest in automated ethics, or the development 
of machines that can perform ethical reasoning. Prior work in automated ethics rarely engages with
philosophical literature on ethics, despite its relevance to the development of AI agents that can
responsibly navigate the world. If automated ethics draws on sophisticated philosophical literature, the ethical reasoning
underlying its decisions will be more nuanced, precise, and trustworthy because it will
take advantage of philosophical progress in developing consistent, mature ethical theories. However, 
faithfully translating complex ethical theories from natural language to the rigid syntax of a computer
program poses technical and philosophical challenges.

### 2022-03-09 paper/output/document/thesis_1_intro.tex
making my way dowtown, typing fast, appendices pass

Machine ethicists recognize the need for automated ethics and have made both theoretical 
(\citep{moralmachineonline}, \citep{davenport}, \citep{moralmachine}, \citep{gabriel}) and practical progress 
(\citep{logicprogramming}, \citep{biology}, \citep{delphi}, \citep{winfield}) towards automating ethics. 
However, prior work in machine ethics using popular ethical theories like deontology (\citep{deon2}, \citep{deon1}), 
consequentialism (\citep{util1}, \citep{util2}, \citep{cloos}), and virtue ethics \citep{berberich} rarely 
engages with philosophical literature and thus often misses philosophers' insights. Even the above example of 
the malfunctioning self-driving car is an instance of Phillipa Foot's trolley problem, 
in which a bystander watching a runaway trolley can pull a lever to kill one instead of three \citep{foot}. 
Decades of philosophical debate have developed ethical theories that can offer nuanced and 
consistent answers to the trolley problem. Like the trolley problem, the moral dilemmas 
that artifical agents face are not entirely new, so solutions to these problems should take advantage of philosophical 
progress. Philosophers are devoted to the creation of better ethical theories, so the 
more faithful that automated ethics is to philosophical literature, the more nuanced, precise, consistent, and
therefore trustworthy it will be.

>>>>

Machine ethicists recognize the need for automated ethics and have made both theoretical 
(\citep{moralmachineonline}, \citep{davenport}, \citep{moralmachine}, \citep{gabriel}) and practical progress 
(\citep{logicprogramming}, \citep{biology}, \citep{delphi}, \citep{winfield}) towards automating ethics. 
However, prior work in machine ethics using popular ethical theories like deontology (\citep{deon2}, \citep{deon1}), 
consequentialism (\citep{util1}, \citep{util2}, \citep{cloos}), and virtue ethics \citep{virtue2} rarely 
engages with philosophical literature and thus often misses philosophers' insights. Even the above example of 
the malfunctioning self-driving car is an instance of Phillipa Foot's trolley problem, 
in which a bystander watching a runaway trolley can pull a lever to kill one instead of three \citep{foot}. 
Decades of philosophical debate have developed ethical theories that can offer nuanced and 
consistent answers to the trolley problem. Like the trolley problem, the moral dilemmas 
that artifical agents face are not entirely new, so solutions to these problems should take advantage of philosophical 
progress. Philosophers are devoted to the creation of better ethical theories, so the 
more faithful that automated ethics is to philosophical literature, the more nuanced, precise, consistent, and
therefore trustworthy it will be.

### 2022-03-09 paper/output/document/thesis_5_discussion.tex
making my way dowtown, typing fast, appendices pass

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents in practice. My work on automating the categorical imperative 
could serve as one component of a partially or fully artificial ethical reasoner, or an ``ethical AI.''
Specifically, my system is a categorical imperative library that takes as input the logical representation 
of a maxim and returns its moral status (if it is obligatory, prohibited, or permissible).
As it stands, my project can evaluate the moral status of maxims represented in my logic and potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{center}
\begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent, which begins with a moral dilemma, 
passes it through an input parser, applies the automated categorical imperative test, and then processes the 
output using an output parser. I contribute the automated categorical imperative component.} \label{fig:AIengine}
\end{figure}
\end{center}
%
\begin{isamarkuptext}%
In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work would need to develop such input and output parsers.

@@

While this account of the formulation of a maxim describes how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. In order for an automated
ethical agent to use the categorical imperative to its fullest extent, the input maxim must be a good-faith 
attempt to capture the agent's principle
of action. However an action is turned into a maxim for my system to process, whether manually as I did
in Chapter \ref{applications} or using an automatic input parser, this transformation must be a good-faith 
attempt to capture the principle of action. Translating everyday situations into appropriate maxims is 
the bulk of the work that a Kantian human 
being does when making decisions. Common misconceptions about Kantian ethics\footnote{For example, critics 
of Kantian ethics worry that the maxim, ``When I am a
man, I will marry a man because I want to spend my life with him'' fails the universalizability
test because if all men only married men, sexual reproduction would stop. This argument implies 
that Kantian ethics is homophobic. Kantians often respond by arguing that the correct formulation of 
this maxim is, ``When I love a man, I will marry him because I want to spend my life with him,'' which
is universalizable because if everyone marries who they love, some men will marry women and others will
marry men.} often result from incorrectly formulated maxims, 
and the entire field of applied Kantian ethics is devoted to generating the right kinds of maxims to test.

This representational question is one of the biggest challenges to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser. Once an AI agent stumbles onto an ethical dilemma, a human being could take over, formulate 
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. For proponents of the ``human-in-the-loop'' model of AI ethics, in which 
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
The outcome of the universalizability test will depend on how the human formulates the maxim; if the 
human puts garbage into the test, the test will return garbage out.

@@

My system's output could be converted into some actionable, useful response with an output parser, 
and then passed back to the AI agent. For example, if the AI agent is equipped to evaluate natural 
language prescriptions, the status of the maxim could be parsed into a natural language sentence. The 
input parser, categorical imperative library, and output parser together constitute an ``ethics engine'' 
that AI developers could use as a black box implementation of an ethical theory.

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent.
An automated version of the categorical imperative could become part of an ethics engine 
for an AI agent, with additional work to parse the input and the output. Effectively, the kind 
of automated ethics I implement could be a library that AI developers use to give AI agents the capacity for 
sophisticated ethical reasoning faithful to philosophical literature. This represents an improvement 
over existing AI ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In this section, I share an example of the kind of philosophical insight that computational ethics 
can prompt and analyze the value that this tool can offer to philosophers.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$. The insight that the FUL should not apply to badly-formed maxims has 
philosophical value and serves as evidence of the potential of computational ethics.

@@

Consider the example badly-formed maxim, ``When eating breakfast, I will eat breakfast in order to 
eat breakfast.'' There is something empty about this maxim because acting on it could never result in 
any action. If an agent adopts this maxim, 
they decide that, in the circumstances ``eating breakfast'' they will perform the act ``eating breakfast''
for the purpose ``eating breakfast.'' In these circumstances, the act has 
already been performed. Adopting this maxim as a law to live by does not change how you live. If you adopt 
this maxim, then when you are eating breakfast, you eat breakfast, but this statement is already tautologically true.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for a 
philosophical debate to show that computational ethics can result in insights with real 
philosophical weight. I use badly formed maxims to resolve the tension between self-doubt and self-respect. 
This debate originated in epistemology, which values doubting your beliefs but also rationally requires
that you believe that you are not mistaken (else you should update your beliefs). 
Self-doubt and self-respect also compete in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves, and it becomes clear that both concepts can
co-exist, as we intuitively desire. I sketch the details of this argument in the rest of this section.

Another reading of the dancing maxim understands it as an expression of self-doubt. Under this 
reading, the question ``When I am dancing, should I be dancing for the sake of dancing?'' is the agent asking, 
``Am I doing the right thing right now?'' Unlike the drowning example, the agent is not asking about the 
next moment, but is expressing doubt about the moral validity of their behavior at this current moment. 
Surely self-doubt does not always undermine the will—after all, self-doubt plays an 
important role in moral reasoning and is often the mark of a thoughtful agent. I argue 
that questions of self-doubt do not actually involve badly-formed maxims, for these are not the maxims 
that the agent is doubting. I first frame this debate by introducing the tension between self-doubt and 
self-respect in epistemology, then explain the parallel tension in ethics, and finally present a 
resolution of this tension.

@@

This debate between self-respect and self-doubt in epistemology can be extended to ethics. When we 
commit ourselves to acting, we cannot simultaneously doubt the validity of our action. If human 
behavior is purposive, then the very act of committing implies that one has sufficient reasons for 
committing. These reasons may be flawed, but in making the commitment, the reasoner accepts them. It 
is contradictory to claim that someone commits and questions simultaneously. Either the commitment 
is not real, or the question is not. I will call the principle that one cannot will a maxim and 
simultaneously question if they should will that maxim ``ethical self-respect'' or ESR.

@@

To resolve this issue, I turn to another interpretation of ethical self-doubt. Under this interpretation, 
when Sara asks, ``Should I really be dancing right now?'' she wants to know if the maxim that 
resulted in the current moment when she is on the dance floor was the right thing to will. She is 
asking if she made the right decision in the past, when she decided to dance. The maxim that initiated 
the dancing may be something like ``When at a wedding, dance for the sake of dancing.'' This is the maxim 
that she is currently acting on, not the badly-formed maxim  from above. Under this interpretation, 
there is no tension at all between self-doubt and self-respect. It is perfectly valid for a reasoner 
to doubt their prior moral judgements, just as it is perfectly rational for a believer to doubt their 
past beliefs \cite[3-4]{christensen}. Such doubt does not undermine the reasoner's decision-making 
capacity and is thus perfectly consistent with ethical self-respect. Moreover, this is a much more 
intuitive question to be asking \emph{because} it is a well-formed maxim. Not only is the badly-formed
interpretation of this maxim problematic for ESR, it is also not the kind of question that is useful for a practical
reasoner to ask, as argued above.

@@

By recognizing that the naive version of ethical self-doubt is a badly formed maxim, I realized 
that there is something wrong with its formulation and modified it to resolve the above debate. 
This demonstrates a larger meta-pattern in Kantian reasoning: many debates in Kantian philosophical
literature revolve around incorrectly formulated maxims, so the insight about badly
formed maxims is significant. Much of the work of a Kantian ethicist is formulating an
appropriate maxim, and badly-formed maxims define one boundary condition for this task. Just as my isolation
of the wrongness of lying in Section \ref{joking} could help guide the formation of morally worthy maxims, 
so can the insight about badly-formed maxims.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{An Argument For Computational Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The insight above is an example of the kind of philosophical progress that can be 
made using computational tools and serves as evidence for the power of computational ethics. The 
idea that the FUL can only hold for well-formed maxims would have been
incredibly difficult to discover without a computer. I discovered it while formulating the FUL because 
Isabelle's proof-finding tools look for edge cases like badly-formed maxims. Badly-formed maxims are 
interesting because they are the kind of thing that is usually ignored in ordinary philosophical inquiry. 
Philosophers usually assume that we are not discussing badly-formed maxims because, as argued above, 
they are not the kind of thing that is immediately relevant to ethics. Computational tools like Isabelle
require that assumptions like the exclusion of well-formed maxims are made precise, and thus force 
philosophers to understand their arguments in a new way. This example demonstrates the contributions
that computational ethics makes: it can quickly check edge cases and it requires that all assumptions be
made explicit and precise.

I do not argue that computational ethics uncovers philosophical insights that humans are incapable 
of reaching. Instead, I claim that insights like the one about the well-formed maxim are much easier
to reach with computational tools. Computational 
tools prompt philosophers to ask new questions that 
lead to insights, and can thus serve as another tool in a philosopher's arsenal, like a 
thought experiment or counterexample.\footnote{Even if computational ethics is possible, some may 
worry that it sacrifices something important about philosophical inquiry. I discuss this concern in Appendix 
\ref{academicphilosophy}.}

@@

Computational ethics is at its infancy. The use of theorem provers in mathematics is just now beginning 
to make headway \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers for ethics occurred in the last few years. The 
fact that this nascent technology is already helping humans reach non-trivial philosophical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

@@

%:%50=20%:%
%:%51=21%:%

@@

%:%60=31%:%
%:%61=32%:%

@@

%:%172=142%:%
%:%173=143%:%
%:%174=144%:%
%:%183=146%:%

@@

%:%197=150%:%
%:%198=151%:%
%:%199=152%:%
%:%208=154%:%

@@

%:%400=336%:%
%:%401=337%:%
%:%402=338%:%
%:%403=339%:%
%:%404=340%:%
%:%405=341%:%
%:%406=342%:%
%:%415=346%:%

@@

%:%489=410%:%
%:%490=411%:%
%:%491=412%:%
%:%500=414%:%

@@

%:%588=492%:%
%:%589=493%:%
%:%590=494%:%
%:%599=497%:%

@@

%:%715=604%:%
%:%716=605%:%
%:%717=606%:%

@@

%:%723=611%:%
%:%724=612%:%
%:%725=613%:%

@@

%:%740=626%:%
%:%741=627%:%
%:%742=628%:%
%:%751=632%:%

@@

%:%812=683%:%
%:%813=684%:%
%:%814=685%:%
%:%823=687%:%

@@

%:%871=725%:%
%:%872=726%:%
%:%873=727%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that takes 
as input the logical representation of a maxim and returns its moral status (if it is obligatory, 
prohibited, or permissible). My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{center}
\begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent, which begins with a moral dilemma, 
passes it through an input parser, applies the automated categorical imperative test, and then processes the 
output using an output parser. I contribute the automated categorical imperative component.} \label{fig:AIengine}
\end{figure}
\end{center}
%
\begin{isamarkuptext}%
In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work would need to develop such input and output parsers.

@@

While the above account how a well-intentioned human agent can determine 
morally relevant circumstances, the challenge remains open for automated ethics. In order for an automated
ethical agent to use the categorical imperative to its fullest extent, the input maxim must be a good-faith 
attempt to capture the agent's principle
of action. However an action is turned into a maxim for my system to process, whether manually as I did
in Chapter \ref{applications} or using an automatic input parser, this transformation must be a good-faith 
attempt to capture the principle of action. Translating everyday situations into appropriate maxims is 
the bulk of the work that a Kantian human 
being does when making decisions. Common misconceptions about Kantian ethics\footnote{For example, critics 
of Kantian ethics worry that the maxim, ``When I am a
man, I will marry a man because I want to spend my life with him'' fails the universalizability
test because if all men only married men, sexual reproduction would stop. This argument implies 
that Kantian ethics is homophobic. Kantians often respond by arguing that the correct formulation of 
this maxim is, ``When I love a man, I will marry him because I want to spend my life with him,'' which
is universalizable because if everyone marries who they love, some men will marry women and others will
marry men.} often result from incorrectly formulated maxims, 
and the entire field of applied Kantian ethics is devoted to generating the right kinds of maxims to test.

The challenge of formulating a maxim is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser. Once an AI agent stumbles onto an ethical dilemma, a human being could take over, formulate 
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. For proponents of the ``human-in-the-loop'' model of AI ethics, in which 
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
The outcome of the universalizability test will depend on how the human formulates the maxim; if the 
human puts garbage into the test, the test will return garbage out.

@@

My system's output could be converted into some actionable, useful response with an output parser, 
and then passed back to the AI agent. For example, if the AI agent is equipped to evaluate natural 
language prescriptions, the status of the maxim could be parsed into a natural language sentence. The 
input parser, categorical imperative, and output parser together constitute an ``ethics engine'' 
that AI developers could use in a variety of AI systems.

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent,
with additional work to parse the input and output of my implementation of the categorical imperative.
Effectively, the kind of automated ethics I implement could be part of a library that AI developers use to 
give AI agents the capacity for sophisticated ethical reasoning faithful to philosophical literature. 
This represents an improvement over existing AI ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In this section, I share an example of the kind of philosophical insight that computational ethics 
can prompt and analyze the value that this tool can offer to philosophers.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$. The insight that the FUL should not apply to badly-formed maxims has 
philosophical value and serves as evidence of the potential of computational ethics.

@@

Consider the example badly-formed maxim, ``When eating breakfast, I will eat breakfast in order to 
eat breakfast.'' There is something empty about this maxim because acting on it could never result in 
any action. If I adopts this maxim as a principle to live by,
I decide that, in the circumstances ``eating breakfast'' I will perform the act ``eating breakfast''
for the purpose ``eating breakfast.'' In these circumstances, the act has 
already been performed. Adopting this maxim as a law to live by does not change how I live. If I adopt 
this maxim, then when I am eating breakfast, I eat breakfast, but this statement is already tautologically true.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for a 
philosophical debate to show that computational ethics can result in insights with real 
philosophical weight. I use badly formed maxims to resolve the tension between self-doubt and self-respect. 
This debate originated in epistemology, which values doubting your beliefs but also rationally requires
that you believe that you are not mistaken (else you should update your beliefs). 
Self-doubt and self-respect also compete in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves, and it becomes clear that both concepts can
co-exist. I sketch the details of this argument in the rest of this section. I will conclude that 
badly-formed maxims serve as one guide to the formulation of maxims, and thus may have bearing on 
other debates in Kantian ethics as well. I first introduce the tension between self-doubt and self-respect 
in epistemology, then explain the parallel tension in ethics, and finally present a resolution of this 
tension using badly-formed maxims.

@@

This debate between self-respect and self-doubt in epistemology can be extended to ethics. When we 
commit ourselves to acting, we cannot simultaneously doubt the validity of our action, just as we cannot
rationally disapprove of our own beliefs. If human 
behavior is purposive, then the very act of committing implies that one has sufficient reasons for 
committing. These reasons may be flawed, but in making the commitment, the reasoner accepts them—that is 
just what it means to commit oneself to acting. It 
is contradictory to claim that someone commits and questions simultaneously. Either the commitment 
is not real, or the question is not. I will call the principle that one cannot will a maxim and 
simultaneously question if they should will that maxim ``ethical self-respect'' or ESR.

@@

If ethical self-doubt as badly-formed maxim contradicts self-respect, then resolving this issue requires a 
different interpretation of ethical self-doubt. Under this interpretation, 
when Sara asks, ``Should I really be dancing right now?'' she wants to know if the maxim that 
resulted in the current moment of dancing was the right thing to will. She is 
asking if she made the right decision in the past, when she decided to dance. The maxim that initiated 
the dancing may be something like ``When at a wedding, dance for the sake of dancing.'' This is the maxim 
that she is currently acting on, not the badly-formed maxim  from above. Under this interpretation, 
there is no tension at all between self-doubt and self-respect. It is perfectly valid for a reasoner 
to doubt their prior moral judgements, just as it is perfectly rational for a believer to doubt their 
past beliefs \cite[3-4]{christensen}. Such doubt does not undermine the reasoner's decision-making 
capacity and is thus perfectly consistent with ethical self-respect. Moreover, this is a much more 
intuitive question to be asking \emph{because} it is a well-formed maxim. Not only is the badly-formed
interpretation of this maxim problematic for ESR, it is also not the kind of question that is useful for a practical
reasoner to ask, as argued above.

@@

By recognizing that the naive version of ethical self-doubt is a badly formed maxim, I realized 
that there is something wrong with its formulation and modified it to resolve the above debate. 
This demonstrates a larger meta-pattern in Kantian reasoning: many debates in Kantian philosophical
literature revolve around incorrectly formulated maxims, which means that the insight about badly
formed maxims may have bearing on these debates. Much of the work of a Kantian ethicist is formulating an
appropriate maxim, and badly-formed maxims define one boundary condition for this task. Just as my isolation
of the wrongness of lying in Section \ref{joking} could help guide the formation of morally worthy maxims, 
so can the insight about badly-formed maxims.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{An Argument For Computational Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The insight above is an example of the kind of philosophical progress that can be 
made using computational tools and serves as evidence for the power of computational ethics. The 
idea that the FUL can only hold for well-formed maxims would have been
incredibly difficult to discover without a computer. I discovered it while formulating the FUL because 
Isabelle's proof-finding tools look for edge cases like badly-formed maxims. Badly-formed maxims are 
interesting precisely because they are the kind of thing that is usually ignored in ordinary philosophical inquiry. 
Philosophers usually assume that we are not discussing badly-formed maxims because, as argued above, 
they are not the kind of thing that is immediately relevant to ethics. Computational tools like Isabelle
require that assumptions like the exclusion of well-formed maxims are made precise, and thus force 
philosophers to understand their arguments in a new way. This precision is also evident in the philosophical 
implications of my examination of lying and joking in Section \ref{joking}, where I isolated the 
specific cause of lying's wrongness. These examples demonstrate the contributions
that computational ethics makes: it can quickly check edge cases and it uncovers imprecise, ambiguous,
or implicit assumptions.

I do not argue that computational ethics uncovers philosophical insights that humans are incapable 
of reaching. Instead, I claim that insights like the one about well-formed maxims are much easier
to reach with computational tools. Computational 
tools prompt philosophers to ask new questions that 
lead to insights, and can thus serve as another tool in a philosopher's arsenal, like a 
thought experiment or counterexample.\footnote{Even if computational ethics is possible, some may 
worry that it sacrifices something important about philosophical inquiry. I discuss this concern in Appendix 
\ref{academicphilosophy}.}

@@

Computational ethics is at its infancy. The use of theorem provers in mathematics is just now beginning 
to make headway \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers for progress in any kind of philosophy occurred 
in the last few years \color{red} and, at the time of this writing, no known work has used theorem 
provers for computational ethics \color{black}. The 
fact that this nascent technology is already helping humans reach non-trivial philosophical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

@@

%:%52=23%:%
%:%53=24%:%

@@

%:%62=32%:%
%:%63=33%:%

@@

%:%180=143%:%
%:%192=145%:%
%:%193=146%:%
%:%194=147%:%

@@

%:%205=151%:%
%:%217=153%:%
%:%218=154%:%
%:%219=155%:%

@@

%:%408=339%:%
%:%420=341%:%
%:%421=342%:%
%:%422=343%:%
%:%423=344%:%
%:%424=345%:%
%:%425=346%:%
%:%426=347%:%

@@

%:%497=411%:%
%:%509=413%:%
%:%510=414%:%
%:%511=415%:%

@@

%:%596=494%:%
%:%608=497%:%
%:%609=498%:%
%:%610=499%:%

@@

%:%717=605%:%
%:%718=606%:%
%:%719=607%:%

@@

%:%725=611%:%
%:%726=612%:%
%:%727=613%:%

@@

%:%748=629%:%
%:%760=631%:%
%:%761=632%:%
%:%762=633%:%

@@

%:%820=684%:%
%:%832=686%:%
%:%833=687%:%
%:%834=688%:%

@@
